\documentclass{article}

\usepackage{array}
\usepackage{graphicx}

\begin{document}

\title{Finding Magnetic Fields, Problems}
\author{GSI: Caleb Eades}
\date{11/13}
\maketitle

\section{Finding Magnetic Fields: Ampere's Law and the Biot-Savart Law}

\subsection{Plane of Current}

A flat sheet of metal of thickness $t$ carries a uniform current density $\vec{j}$.
\begin{itemize}
\item[(a)] Using Biot-Savart's Law, calculate the magnitude of the $\vec{B}$ field created at a point a distance $y$ above the sheet, considering the infinite plane as a succession of infinite wires.
\item[(b)] Recalculate $\vec{B}$ using a faster method.
\end{itemize}

\textit{(Source: Bordel Fall 2013 Final, Problem 5)}

\subsection{Loops and Wires}

A circular loop of radius $R$ carries a steady current $I_2$ in a clockwise direction as shown on the top figure. The center of the loop is a distance $d>R$ above a long straight wire carrying current $I_1$.
\begin{itemize}
\item[(1)] What is the direction and magnitude of the steady current $I_1$ in the wire if the magnetic field at the center of the loop is zero?
\item[(2)] Is the force between the ring and the wire zero, repulsive or attractive?
\end{itemize}

\textit{(Source: Corsini Spring 2015 Final, Problem 7)}

\subsection{Bent Wire}

A long wire is bent into the hairpin-like shape.
\begin{itemize}
\item[(a)] What is the direction of the magnetic field at the center of the half-circle of radius $d$?
\item[(b)] What is the magnitude of the magnetic field at that point?
\item[(c)] Suppose an electron is at point P and has a velocity $v$ pointing into the page. What force does the magnetic field exert on it?
\end{itemize}

\textit{(Source: Lanzara Fall 2014 Final, Problem 2)}

\subsection{Spinning Charges}

A hollow sphere of radius $R$ and total charge $Q$ is spinning along the z-axis with angular velocity $\omega$. What is the magnetic field at the center of the sphere (magnitude and direction)? Express it in terms of $Q$, $R$, $\omega$, and $\mu_0$.

\textit{(Source: Speliotopoulos Fall 2014 Final, Problem 3)}

\subsection{Arc-ed Wire}

A section of wire with current $I$ running through it has an opening of angle $\theta_0$ and radius $a$. What is the magnetic field (direction and magnitude) at the center of the opening?

\textit{(Source: Speliotopoulos Spring 2014 Final, Problem 2)}


\end{document}
