\documentclass{article}

\usepackage{array}
\usepackage{amsmath}

\begin{document}

\title{Week 6, Session 2 Solutions}
\author{GSI: Caleb Eades}
\date{9/27}
\maketitle

\section{Coulomb's Law}

\subsection{Charges in a Bowl}

\begin{itemize}
	\item[(a)] If the charges are separated by some angle $\theta$, then the physical distance separating them is
	\begin{equation}
	d = 2Rsin\left(\frac{\theta}{2}\right)
	\end{equation}
	From Newton's second law, we garner
	\begin{align*}
	Ncos\frac{\theta}{2} &= mg \\
	Nsin\frac{\theta}{2} &= \frac{1}{4\pi\epsilon_0}\frac{Q^2}{d^2} \\
	&= \frac{1}{4\pi\epsilon_0}\frac{Q^2}{4R^2sin^2\frac{\theta}{2}}
	\end{align*}
	From the first equation, $N = \frac{mg}{cos\frac{\theta}{2}}$. 		From the second,
	\begin{align*}
	Q^2 &= 16\pi\epsilon_0 R^2 N sin^3\frac{\theta}{2} \\
	&= 16\pi\epsilon_0 mgR^2 tan\frac{\theta}{2}sin^2\frac{\theta}{2} \\
	Q &= 4Rsin\frac{\theta}{2}\sqrt{\pi\epsilon_0 mgtan\frac{\theta}{2}
	\end{align*}
	\item[(b)] Is this equilibrium stable? If so, calculate the frequency of small oscillations about it.
\end{itemize}

\subsection{Oh Charge, Where Art Thou?}

Two point charges are located on the $x$ axis. They are both positive, bu the one located at $x=0$ has a charge of $q$ while the one located at $x=L$ has a charge of $4q$. If a third charge is placed on the $x$ axis in between the two charges so that the net force on ANY of the charges is zero, determine the magnitude of the third charge and its location.

(\textit{Source: physics-prep.com})

\subsection{Return of the Spring}

A spring with spring constant $k_s$ and rest length $L$ has positive charges $Q$ attached to either end.
\begin{itemize}
	\item[(a)] Find an equation that will determine the length $D$ of the spring, once the charges have come to rest.
	\item[(b)] Repeat part (a), this time assuming that the charges on either end are both \textit{negative}.
	\item[(c)] Repeat again, this time assuming that the charges on either end have \textit{opposite} signs.
\end{itemize}

(\textit{Source: workbook })

\subsection{A Balancing Act}

A charge $q$ hangs on the end of a string while another charge $-Q$ of mass $m$ is brought beneath it.
\begin{itemize}
	\item[(a)] At what distance $d$ below the hanging mass is the charge $-Q$ in equilibrium?
	\item[(b)] Is this equilibrium stable? If so, find the frequency of small oscillations about it.
\end{itemize}

\subsection{Dipoles}

(Challenge) Find the electric field due to a dipole located at the origin both along its axis and in the plane perpendicular to its axis. (Hint: find the field from two point charges with charge $q$ located at $y=d/2$ and charge $-q$ located at $y=-d/2$ and then take the limit as $d$ goes to zero.)

\end{document}
